### 2024-10-24 paper/paper.qmd
Organized the paper section structure

---
title: "Counts of Crimes and Their Subtypes in Toronto from 2014 to 2023"
subtitle: "Details and Trends"
author: 
  - Justin (Jiazhou) Bi
thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/reported-crime/tree/master."
date: Sept. 23, 2024
date-format: long
abstract: "This report analyzes crime trends and subtypes in Toronto from 2014 to 2023, highlighting fluctuations and speculating possible socio-economic, technological, and pandemic-related factors influencing criminal activity. Using data from Toronto Open Data, the study aggregates and examines reported crimes in categories such as property crimes, violent offenses, and cybercrime. The results show a significant increase in reported crimes, from 113,426 in 2014 to 169,620 in 2023. A notable drop in crime during 2020 and 2021 corresponds with COVID-19 lockdowns, which reduced certain physical crimes while increasing cybercrimes like fraud. Despite the overall rise in crime, clearance rates for crimes like fraud and theft significantly declined, underscoring challenges for law enforcement."
format: pdf
jupyter: python3 
number-sections: true
bibliography: references.bib
---

# Introduction

Crime is defined as "an action or omission which constitutes an offense and is punishable by law." (@TheOxfordDictionary_1992) However, this seemingly straightforward definition can misguide us to think that crime is a simple and clear concept with specific defining criteria, a swift process of judgment, and pre-defined punishment. Crime is a complete concept that affects every aspect of all individuals, communities, and societies. Crimes can reshape the power dynamics of our social structure, influence our public policies, and even change our day-to-day interactions with others. As a result, understanding the trends in crimes and the variations in their subtypes is crucial for policymakers, law enforcement, and every member of our society to devise better strategies to prevent crimes from happening. Therefore, this report aims to explore and examine the current of reported crimes in Toronto from 2014 to 2023, discover and highlight some fluctuations, and propose some potential underlying factors that might have contributed to these trends. The raw data being used in the report is available from Toronto Open Data. (@TorontoOpenData) For the data analysis, I have used pandas (@pandas). For data visualization, I have used Seaborn. (@seaborn) All the tables were generated by using Plotly. (@plotly)

This analysis can be helpful by providing valuable insights into current crime activity trends. Certain crimes, such as violent offenses, theft, and property crimes, are believed to be highly correlated with broader socio-economic conditions. (@lofstrom2016crime) Meanwhile, with new technology emerging, other types of criminal activity, such as technology-facilitated fraud, are becoming more prominent. According to RCMP's report, Canadian victims have lost 530 million dollars due to fraud. Thus, examining various types of crime can assist us to better understand these crimes and devise better strategies to better protect our safety and properties.

On the other hand, between 2014 and 2023, the world experienced significant transformations that may impact crime rates. In particular, the COVID-19 pandemic, which began in 2020 and ended in early 2022, has disrupted our daily routines and led to shifts in criminal behaviour. Lockdowns and other public health measures reduced opportunities for certain types of crime, such as burglaries and public disturbances, while simultaneously leading to increases in domestic violence and cybercrime. Additionally, economic hardship caused by the pandemic, including rising unemployment rates and financial instability, may have driven a surge in financially motivated crimes, such as fraud and theft.

This report will delve into crime statistics from 2014 to 2023 in Toronto. I will closely examine all the reported crime categories and specific subtypes of crimes. By analyzing these trends, I aim to clarify how crime has evolved over the last decade and identify possible socio-economic, technological, and legal factors that have driven these changes. Through this analysis, we can gain valuable insights into how best to address the challenges posed by both traditional and emerging forms of crime in the future.

# Data

The dataset used here is downloaded from the Toronto Open Data website via @TorontoOpenData. This dataset contains all the reported crimes that happened in Toronto from 2014 to 2023. This dataset is grouped by the year of the reported crime, its category and belonging subtype, and the count of the subtype being reported and cleared for that year for each division. Because I am examining the crime pattern in the city, I have dropped the division information and aggregated the existing data according to their subtype and the year of the crime being reported. In the following subsections, I will review all the variables used in this report and provide some basic descriptive statistics. The first five rows of the cleaned data used for analysis are attached (@tbl-table1).

```{python}
#| label: tbl-table1
#| tbl-cap: Example of Cleaned Data
#| warning: false
#| echo: false

import pandas as pd
data = pd.read_csv('../data/analysis_data/analysis_data.csv', index_col=False)
data.index = data.index + 1
data.head()
# styled_data = data.head().style.format({'ClearRate': "{:.2f}"}).hide().set_table_styles([{"selector":"th.col_heading,td","props":[("width","100px")]}])
# styled_data
#data.values[:5]
#print(data.head().to_string(index=False))

```

## Report Year
The report year variable is the number of crimes being reported. In this dataset, the data spans from 2014 to 2023, encompassing ten years. No month or date information was given; thus, there are only ten different values for this variable in chronological order.

## Category
The category includes information about the nature of the crime. There are six crime categories: Crimes Against Property, Crimes Against the Person, Other Federal Statute Violations, Other Criminal Code Violations, Controlled Drugs and Substances Act, and Criminal Code Traffic. They are listed in the table below (@tbl-table2).

```{python}

#| echo: false
#| label: tbl-table2
#| warning: false
#| tbl-cap: Six Crime Categories
import pandas as pd
import plotly.graph_objects as go
data = pd.read_csv('../data/analysis_data/analysis_data.csv')
unique_category = data.Category.unique()
category_df = pd.DataFrame(unique_category, columns=['Categories'])
category_df.index = category_df.index + 1
category_df
```

## Subtype
There exist multiple subtypes under each crime category. The following is an exhaustive table (@tbl-table3) of all crimes' subtypes and their respective category.

```{python}
#| echo: false
#| label: tbl-table3
#| warning: false
#| tbl-cap: Crime Categories and Subtypes
import pandas as pd
data = pd.read_csv('../data/analysis_data/analysis_data.csv')
unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
unique_category_subtype_df.reset_index(drop=True, inplace=True)
unique_category_subtype_df.index = unique_category_subtype_df.index + 1
unique_category_subtype_df
```

## Count
In the original table, this value is grouped by the subtype of the crime, the division, and the year when the crime was reported. The original count indicates the number of a specific subtype of crime reported within a particular division for the year. However, as mentioned before, because I am only interested in all the crimes in the City of Toronto, I have dropped the division information and aggregated the count from all the divisions to a single value. Therefore, for each subtype of the crimes, a total count of that subtype is reported in a single year.

## Count_Cleared
These are the counts of crimes identified as cleared. In plain words, these are crimes that are dealt/solved. I have taken the same approach for this column as the previous one. After cleaning the data,f, or each subtype of the crimes, there is a total count of that subtype being reported that is also cleared in a single year.

## Case_Clearing_Rate
This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes. A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime. The value is ranged from 0 to 100%.

# Model

Our modeling strategy aims to explore the relationship between occupation, age, gender, education, and income. To achieve this, we implemented three models: Linear Regression with Interaction Terms, Random Forest, and XGBoost, each selected based on their specific strengths for predictive accuracy and interpretability.

## Linear Regression Model with Interaction Term

## Random Forest

## XGBoost

# Results

# Discussion

## Weaknesses and next steps

\newpage

# Appendix

## Data cleaning

## Model Details

# References

### 2024-10-24 paper/paper.qmd
Abstract and Introduction updated

@@
 ---
-title: "Counts of Crimes and Their Subtypes in Toronto from 2014 to 2023"
-subtitle: "Details and Trends"
+title: "Personal Total Income of Individual in the United States of America in 2023"
+subtitle: "A Predictive Model Using the American Community Survey Data"
 author: 
-  - Justin (Jiazhou) Bi
-thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/reported-crime/tree/master."
-date: Sept. 23, 2024
+  - Justin (Jiazhou) Bi and Weiyang Li
+thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/ACS-Income-Model/tree/main."
+date: Oct 25, 2024
 date-format: long
-abstract: "This report analyzes crime trends and subtypes in Toronto from 2014 to 2023, highlighting fluctuations and speculating possible socio-economic, technological, and pandemic-related factors influencing criminal activity.
-Using data from Toronto Open Data, the study aggregates and examines reported crimes in categories such as property crimes, violent offenses, and cybercrime.
-The results show a significant increase in reported crimes, from 113,426 in 2014 to 169,620 in 2023.
-A notable drop in crime during 2020 and 2021 corresponds with COVID-19 lockdowns, which reduced certain physical crimes while increasing cybercrimes like fraud.
-Despite the overall rise in crime, clearance rates for crimes like fraud and theft significantly declined, underscoring challenges for law enforcement."
+abstract: "This project aims to build a predictive model for personal total income using various demographic attributes, such as marital status, occupation, and residing state.
+We explored three popular machine learning algorithms: Random Forest, Linear Regression, and Extreme Gradient Boosting.
+Ultimately, Extreme Gradient Boosting was selected due to its superior predictive performance.
+While the model achieved a high R-squared value, it also exhibited a high Mean Squared Error, indicating challenges in accurately predicting all data points.
+Future steps for improving the model may involve incorporating additional demographic features to enhance model accuracy and reduce errors.
+Additionally, fine-tuning hyperparameters, experimenting with feature engineering, and handling outliers could further improve performance."
+
 format: pdf
@@
 
-Crime is defined as "an action or omission which constitutes an offense and is punishable by law." (@TheOxfordDictionary_1992) However, this seemingly straightforward definition can misguide us to think that crime is a simple and clear concept with specific defining criteria, a swift process of judgment, and pre-defined punishment.
-Crime is a complete concept that affects every aspect of all individuals, communities, and societies.
-Crimes can reshape the power dynamics of our social structure, influence our public policies, and even change our day-to-day interactions with others.
-As a result, understanding the trends in crimes and the variations in their subtypes is crucial for policymakers, law enforcement, and every member of our society to devise better strategies to prevent crimes from happening.
-Therefore, this report aims to explore and examine the current of reported crimes in Toronto from 2014 to 2023, discover and highlight some fluctuations, and propose some potential underlying factors that might have contributed to these trends.
-The raw data being used in the report is available from Toronto Open Data.
-(@TorontoOpenData) For the data analysis, I have used pandas (@pandas).
-For data visualization, I have used Seaborn.
-(@seaborn) All the tables were generated by using Plotly.
+Income is a fundamental concept as it affects nearly every member of society, influencing access to resources, quality of life, and overall economic stability.
+At the same time, it is also a highly complex concept to define, as it can be interpreted differently depending on various perspectives, such as economic, social, legal, and tax purposes.
+Therefore, we must examine various demographical factors that may impact income to understand its concept better.
+Our project defines income as "all forms of net financial resources generated or lost from work, investments, or other activities." This project aims to predict personal income using data from the 2023 American Community Survey (ACS) (@ipums_usa_v15).
+The ACS provides detailed demographic information, including age, education, employment, marital status, mortgage status, veteran status, etc...
+As a result, it provides a robust dataset for studying potential factors that may influence income.
+
+Due to the timeframe limitations, the findings of this project are specifically applicable to the 2023 dataset.
+As such, the results may vary across different periods, and their generalizability has not been assessed within this project.
+The raw data used in this analysis was sourced from IPUMS using the IPUMS API.
+(@ipums_api) We utilized pandas (@pandas) for data processing, and data visualization was carried out using Seaborn.
+(@seaborn) Additionally, all tables presented in this report were generated with Plotly.
 (@plotly)
 
-This analysis can be helpful by providing valuable insights into current crime activity trends.
-Certain crimes, such as violent offenses, theft, and property crimes, are believed to be highly correlated with broader socio-economic conditions.
-(@lofstrom2016crime) Meanwhile, with new technology emerging, other types of criminal activity, such as technology-facilitated fraud, are becoming more prominent.
-According to RCMP's report, Canadian victims have lost 530 million dollars due to fraud.
-Thus, examining various types of crime can assist us to better understand these crimes and devise better strategies to better protect our safety and properties.
+Income inequality is a serious problem and can lead to significant consequences and it may affect social structure, economic stability, as well as policy-making processes.
+Researchers found that demographic factors, such as education level and occupation type, can significantly impact an individual's financial standing.
+(@Autor_2013) Similarly, social relationships, such as marital status and household characteristics, affect income dynamics.
+(@WaiteGallagher_2000) Therefore, applying machine learning models to analyze the ACS dataset and using the information mentioned to predict personal income is reasonable.
 
-On the other hand, between 2014 and 2023, the world experienced significant transformations that may impact crime rates.
-In particular, the COVID-19 pandemic, which began in 2020 and ended in early 2022, has disrupted our daily routines and led to shifts in criminal behaviour.
-Lockdowns and other public health measures reduced opportunities for certain types of crime, such as burglaries and public disturbances, while simultaneously leading to increases in domestic violence and cybercrime.
-Additionally, economic hardship caused by the pandemic, including rising unemployment rates and financial instability, may have driven a surge in financially motivated crimes, such as fraud and theft.
+We have chosen 2023 ACS dataseet because it offers many demographical details, as well as self-reported total income of the individuals.
+The sample is also representative and covers all geographical locations in the USA.
+Furthermore, it provides easy access to their database through API.
+We will employ advanced predictive models, including random forest, lineaer regression, and extreme gradient boosting machine to estimate personal income based on demographic factors to provide insights into which potential factors have the most significant impact on personal income.
+We hope the findings generated from our project can provide insights to policy-making process so that we can, hopefully, reduce income inequality and support further economic development in the future.
 
-This report will delve into crime statistics from 2014 to 2023 in Toronto.
-I will closely examine all the reported crime categories and specific subtypes of crimes.
-By analyzing these trends, I aim to clarify how crime has evolved over the last decade and identify possible socio-economic, technological, and legal factors that have driven these changes.
-Through this analysis, we can gain valuable insights into how best to address the challenges posed by both traditional and emerging forms of crime in the future.
+If we can successfully predict personal income using demographic information from the survey, this analysis will be able to contribute to a nuanced understanding of the American economy in 2023.
+It can help to reveal how demographic changess, educational attainment, and employment status/type can impact personal income.
+On the other hand, if the model cannot predict personal income with a high degree of confidence, this would suggest that the factors included in the analysiscannot not fully capture the intricacies of economic behavior and individual circumstances.
+As a result, additional variables, such as detailed education background like school of graduation, languages spoken, health status, etc, might be necessary in accurately predicting income.
 
@@
 import pandas as pd
-data = pd.read_csv('../data/analysis_data/analysis_data.csv', index_col=False)
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
 data.index = data.index + 1
@@
 #| tbl-cap: Six Crime Categories
-import pandas as pd
-import plotly.graph_objects as go
-data = pd.read_csv('../data/analysis_data/analysis_data.csv')
-unique_category = data.Category.unique()
-category_df = pd.DataFrame(unique_category, columns=['Categories'])
-category_df.index = category_df.index + 1
-category_df
+# import pandas as pd
+# import plotly.graph_objects as go
+# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
+# unique_category = data.Category.unique()
+# category_df = pd.DataFrame(unique_category, columns=['Categories'])
+# category_df.index = category_df.index + 1
+# category_df
 ```
@@
 #| tbl-cap: Crime Categories and Subtypes
-import pandas as pd
-data = pd.read_csv('../data/analysis_data/analysis_data.csv')
-unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
-unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
-unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
-unique_category_subtype_df.reset_index(drop=True, inplace=True)
-unique_category_subtype_df.index = unique_category_subtype_df.index + 1
-unique_category_subtype_df
+# import pandas as pd
+# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
+# unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
+# unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
+# unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
+# unique_category_subtype_df.reset_index(drop=True, inplace=True)
+# unique_category_subtype_df.index = unique_category_subtype_df.index + 1
+# unique_category_subtype_df
 ```

### 2024-10-25 paper/paper.qmd
Write Analysis for Model section (except insert graphs)

@@
 
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 
-# Data
+# Data {#sec-data}
 
@@
 
-# Model
+# Exploratory Data Analysis {#sec-eda}
 
-Our modeling strategy aims to explore the relationship between occupation, age, gender, education, and income.
-To achieve this, we implemented three models: Linear Regression with Interaction Terms, Random Forest, and XGBoost, each selected based on their specific strengths for predictive accuracy and interpretability.
+# Model {#sec-model}
 
-## Linear Regression Model with Interaction Term
+Our modeling strategy aims to explore the relationship between personal characteristics and income.
+We implemented three models: Linear Regression with Interaction Terms, Random Forest, and XGBoost, each chosen for their unique strengths in prediction and interpretation.
 
-## Random Forest
+Background details are included in Appendix @sec-model-details.
 
-## XGBoost
+## Linear Regression Model with Interaction Term {#sec-linear-regression-model-with-interaction-term}
 
-# Results
+### Model set-up
 
-# Discussion
+We implemented a linear regression model with an interaction term to explore the relationship between personal characteristics and income. The general form of the model is:
+
+$$
+y=\beta_0+\beta_1X_1+\beta_2X_2+...+\beta_nX_n+\beta_{\text{interaction}}(X_i\times X_j)+\epsilon
+$$
+
+where:
+
+-   y represents the total income (`INCTOT`),
+
+-   $X_1, X_2, ... , X_n$ are the predictor variables, including both categorical and numerical features,
+
+-   $X_i \times X_j$ is the interaction term between education and sex (`EDUC_SEX_INTERACTION`)
+
+-   $\beta_0$ is the intercept,
+
+-   $\beta_1, \beta_2, ... , \beta_n, \beta_{\text{interaction}}$ are the coefficients for the respective predictors and interaction term.
+
+To ensure the robustness of our linear regression model, we conducted a series of steps aimed at minimizing multicollinearity and improving model interpretability.
+Initially, we included all relevant predictor variables, both categorical and numerical, to explore the full range of relationships between personal characteristics and income.
+Next, we calculated the Variance Inflation Factor (VIF) for each predictor to identify potential multicollinearity issues.
+Typically, a VIF value greater than 10 indicates problematic multicollinearity, which can distort the model's estimates.
+In our analysis, variables such as `VESTAT` (Veteran Status) and `AGE` had VIF values exceeding 13, suggesting they were highly correlated with other predictors.
+To address this, we removed these high-VIF variables from the model.
+Afterward, we refit the linear regression model using the reduced set of predictors, which resulted in improved performance metrics, including a lower Mean Squared Error and a more interpretable set of coefficients.
+
+Note that according to our previous analysis in @#sec-eda, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
+This allowed us to model the non-linear relationship between these two variables and their joint impact on the outcome variable.
+
+::: {#tbl-coefficients}
+**Table 1: Coefficients from the Linear Regression Model**
+
+| Feature                     | Coefficient |
+|-----------------------------|-------------|
+| `cat__MORTGAGE`             | 9971.56     |
+| `num__AGE`                  | 397.66      |
+| `cat__STATEICP`             | 91.39       |
+| `cat__IND1990`              | -44.49      |
+| `cat__OCC2010`              | -201.61     |
+| `num__EDUC_SEX_INTERACTION` | -2799.42    |
+| `cat__MARST`                | -4929.08    |
+| `cat__GQ`                   | -7172.27    |
+| `cat__SCHLTYPE`             | -14947.02   |
+:::
+
+*Table 1: Linear Regression Coefficients Table.*
+
+We run the model in Python [@cite_python] using the `scikit-learn` package [@cite_sklearn].
+We use the default settings for the `LinearRegression` function from `scikit-learn`.
+We use the `statsmodels` package [@cite_statsmodels] to calculate the VIF for each predictor.
+
+### Model Justification
+
+In evaluating our model's performance, we selected key metrics to assess both its fit and predictive accuracy, including the Variance Inflation Factor (VIF), R-squared ($R^2$), and Mean Squared Error (MSE).
+These metrics were chosen to provide a comprehensive evaluation of the model's robustness and reliability in capturing the relationship between personal characteristics and income.
+
+The VIF was used to diagnose multicollinearity among predictors, helping us ensure that no variables were highly correlated, which could distort the interpretation of the model's coefficients.
+
+::: {#tbl-VIF}
+**Table 2: Variance Inflation Factors (VIF) for Predictor Variables**
+
+| Feature                     | VIF      |
+|-----------------------------|----------|
+| cat\_\_STATEICP             | 3.874599 |
+| cat\_\_GQ                   | 1.002446 |
+| cat\_\_MORTGAGE             | 2.614057 |
+| cat\_\_MARST                | 1.754080 |
+| cat\_\_SCHLTYPE             | 1.242432 |
+| cat\_\_OCC2010              | 2.953611 |
+| cat\_\_IND1990              | 6.566724 |
+| num\_\_AGE                  | 7.373021 |
+| num\_\_EDUC_SEX_INTERACTION | 7.066118 |
+:::
+
+*Table 2: This table displays the Variance Inflation Factors (VIF) for each of the predictor variables in the model. VIF values greater than 10 indicate potential multicollinearity issues.*
+
+We chose $R^2$ as it provides a measure of how well the model explains the variance in the dependent variable, making it an important indicator of the model’s explanatory power.
+The MSE was selected to quantify the average squared difference between observed and predicted values, offering a clear view of the model’s prediction error.
+
+::: {#tbl-modelperformance}
+**Table 3: Model Performance Metrics**
+
+| Metric                         | Value            |
+|--------------------------------|------------------|
+| Mean Squared Error (MSE)       | 7,178,302,129.51 |
+| Root Mean Squared Error (RMSE) | 84,724.86        |
+| R-squared                      | 0.11099          |
+| Adjusted R-squared             | 0.11097          |
+:::
+
+*Table 3: This table presents the performance metrics of the linear regression model, including MSE, RMSE, R-squared, and Adjusted R-squared.
+These metrics are used to assess the model's fit and predictive accuracy.*
+
+Additionally, residual analysis was conducted by plotting residuals against fitted values to check for randomness around zero, ensuring that the model satisfied the assumptions of linearity and homoscedasticity.
+
+We use the statsmodels package [@cite_statsmodels] to calculate the VIF for each predictor.
+We use the matplotlib [@cite_matplotlib] and seaborn [@seaborn] to visualize the residuals.
+We use the scikit-learn [@cite_sklearn] package to calculate the $R^2$ score.
+
+## Random Forest Model {#sec-random-forest}
+
+### Model set-up
+
+To analyze the non-linear relationship between personal characteristics and income, we employed a Random Forest model.
+Random Forest is an ensemble learning method that builds multiple decision trees and averages their predictions, reducing overfitting and capturing complex, non-linear relationships in the data.
+We utilized Bagging (Bootstrap Aggregating) and random feature selection to enhance model performance and reduce overfitting.
+Bagging involves generating multiple bootstrap samples from the original training set, with each sample used to train an independent decision tree.
+By averaging the predictions of multiple trees, Bagging reduces the variance of the model, thereby improving its robustness.
+Additionally, Random Forest introduces further randomness by selecting a subset of features at each split, which reduces correlation among trees and improves the model’s generalization ability.
+
+To optimize the model, we conducted hyperparameter tuning, focusing on two key parameters: `n_estimators` (the number of trees) and `max_depth` (the maximum depth of each tree).
+Increasing `n_estimators` typically decreases model variance and improves stability, but too many trees can result in diminishing returns with increased computation time.
+Initially, we set `n_estimators` to 100, but through Grid Search, we tested values of 100, 200, and 300 and found that 300 trees yielded the best performance.
+Similarly, the `max_depth` parameter controls the complexity of the trees.
+While deeper trees can capture more intricate data patterns, they may also overfit.
+We initially set `max_depth` to 15 and test 10, 15, 20 but found that reducing it to 10 during hyperparameter tuning improved generalization by avoiding overfitting.
+Additionally, we tuned `min_samples_split` and `min_samples_leaf` to prevent overfitting by ensuring that nodes have a sufficient number of samples before splitting.
+
+We run the model in Python [@cite_python] using the scikit-learn package [@cite_sklearn] and pandas [@pandas].
+We used the GridSearchCV function from scikit-learn for hyperparameter tuning and subsampled the training data using pandas.
+The RandomForestRegressor function from scikit-learn was employed for model fitting.
+
+### Model Justification
+
+To assess the performance of our Random Forest model, we evaluated several key metrics, including Mean Squared Error (MSE), Root Mean Squared Error (RMSE), feature importance, and residual analysis.
+These evaluations provided insights into the model’s predictive accuracy, interpretability, and overall fit.
+
+First, we calculated the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) to assess the overall prediction accuracy of the model.
+These metrics provided insight into how well the model captured the variance in income based on the selected features.
+
+::: {#tbl-randomforestMSE}
+**Table 4: Best Random Forest Model Performance Metrics**
+
+| Metric          | Value            |
+|-----------------|------------------|
+| Best Model MSE  | 6,057,654,880.33 |
+| Best Model RMSE | 77,830.94        |
+:::
+
+*Table 4: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the best-tuned Random Forest model, demonstrating the model's performance after hyperparameter tuning.*
+
+Additionally, we performed a feature importance analysis, which allowed us to understand the relative significance of each predictor variable in determining income.
+This is particularly useful for verifying that the model identifies the key factors influencing income as expected.
+
+::: {#tbl-randomforestimportance}
+**Table 5: Feature Importance in the Best Random Forest Model**
+
+| Feature         | Importance |
+|-----------------|------------|
+| cat\_\_EDUC_new | 0.337465   |
+| cat\_\_IND1990  | 0.243701   |
+| num\_\_AGE      | 0.193615   |
+| cat\_\_SEX      | 0.149471   |
+| cat\_\_MARST    | 0.055754   |
+| cat\_\_MORTGAGE | 0.012897   |
+| cat\_\_SCHLTYPE | 0.004205   |
+| cat\_\_VETSTAT  | 0.002891   |
+| cat\_\_OWNERSHP | 0.000000   |
+:::
+
+*Table 5: This table lists the feature importance values for the best-performing Random Forest model.
+The higher the importance value, the more significant the feature is in predicting the target variable (income).*
+
+We also conducted a residual analysis by plotting the residuals (the difference between actual and predicted values) against the predicted values.
+This step helped us evaluate whether the model appropriately captured the underlying patterns in the data, ensuring that there were no significant biases or violations of model assumptions.
+
+The feature importances were extracted from the best model using RandomForestRegressor from scikit-learn [@cite_sklearn] and displayed using pandas [@pandas].
+We use the matplotlib [@cite_matplotlib] to visualize the residuals.
+
+## XGBoost {#sec-xgboost}
+
+### Model set-up
+
+XGBoost (Extreme Gradient Boosting) is a powerful machine learning algorithm that builds an ensemble of decision trees using gradient boosting techniques.
+In gradient boosting, each subsequent tree is trained to correct the residuals of the previous trees, iteratively improving the model’s predictive accuracy.
+XGBoost enhances this process by incorporating regularization techniques to control overfitting, making it an ideal choice for capturing complex, non-linear relationships in data.
+
+To optimize the performance of our XGBoost model, we employed GridSearchCV to conduct hyperparameter tuning.
+We explored a range of hyperparameters to find the optimal configuration for our dataset.
+The grid search included the learning rate (`learning_rata`), max depth (`max_depth`), number of trees (`n_estimators)`, and minimum child weight (`min_child_weight`).
+Specifically, we tested learning rates of 0.01, 0.05, and 0.1, tree depths of 4, 6, and 8, and `n_estimators` values of 100, 200, and 300.
+Using GridSearchCV with a cross-validation fold of 3, we systematically evaluated different combinations of these hyperparameters.
+The model was evaluated using negative Mean Squared Error (MSE) as the scoring metric.
+The optimal model used a learning rate of `0.1`, a maximum tree depth of `6`, a minimum child weight of `5`, and `300` boosting rounds.
+This configuration provided a good balance between model complexity and predictive accuracy, capturing important non-linear relationships while minimizing overfitting.
+
+We implemented the XGBoost model in Python [@cite_python] using the scikit-learn package [@cite_sklearn], with pandas [@pandas] for data processing and xgboost [@cite_xgboost] for model fitting.
+The initial model was trained using default parameters of the XGBRegressor function, which was subsequently fine-tuned using GridSearchCV from scikit-learn [@cite_sklearn] for hyperparameter optimization.
+
+### Model Justification
+
+Similar to the Random Forest model, the XGBoost model was trained on a 70-30 train-test split.
+We evaluated the model's performance using Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) to assess its predictive accuracy.
+
+::: {#tbl-XGBoostMSE}
+**Table 6: XGBoost Model Performance Metrics**
+
+| Metric                         | Value            |
+|--------------------------------|------------------|
+| Mean Squared Error (MSE)       | 5,901,467,894.81 |
+| Root Mean Squared Error (RMSE) | 76,821.01        |
+:::
+
+*Table 6: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the final XGBoost model, reflecting its predictive accuracy.*
+
+We performed a residual analysis by plotting the residuals against the predicted values to assess the model's fit.
+This allowed us to check whether the XGBoost model accurately captured the patterns in the data and to verify that no major biases or violations of model assumptions were present.
+
+We also performed cross-validation to ensure the robustness of the XGBoost model.
+This technique allowed us to evaluate the model’s performance across different subsets of the data, helping to confirm that the model generalizes well and is not overly sensitive to any particular portion of the training data.
+
+::: {#tbl-XGBoostCV}
+**Table 7: XGBoost Cross-Validation Performance**
+
+| Metric                | Value     |
+|-----------------------|-----------|
+| Cross-Validation RMSE | 71,676.11 |
+:::
+
+*Table 7: This table shows the Root Mean Squared Error (RMSE) from cross-validation, demonstrating the model’s performance across different subsets of the data.*
+
+We used matplotlib [@cite_matplotlib] to perform a residual analysis against the predicted values from the XGBoost model.
+We performed cross-validation using cross_val_score function from scikit-learn  [@cite_sklearn] to evaluate the robustness of the model.
+
+# Results {#sec-results}
+
+## Liner Regression Model
+
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 
-## Data cleaning
+## Data Cleaning {#sec-data-cleaning}
 
-## Model Details
+## Model Details {#sec-model-details}
+
+\newpage
 

### 2024-10-25 paper/paper.qmd
Update data importing path & Modify the updated data in Model Section in Paper

@@
 
-Note that according to our previous analysis in @#sec-eda, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
+Note that according to our previous analysis in @sec-eda, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
 This allowed us to model the non-linear relationship between these two variables and their joint impact on the outcome variable.
@@
 |-----------------------------|-------------|
-| `cat__MORTGAGE`             | 9971.56     |
-| `num__AGE`                  | 397.66      |
-| `cat__STATEICP`             | 91.39       |
-| `cat__IND1990`              | -44.49      |
-| `cat__OCC2010`              | -201.61     |
-| `num__EDUC_SEX_INTERACTION` | -2799.42    |
-| `cat__MARST`                | -4929.08    |
-| `cat__GQ`                   | -7172.27    |
-| `cat__SCHLTYPE`             | -14947.02   |
+| `cat__MORTGAGE`             | 11267.90    |
+| `num__AGE`                  | 415.91      |
+| `cat__STATEICP`             | 126.39      |
+| `cat__IND1990`              | -43.18      |
+| `cat__OCC2010`              | -139.55     |
+| `num__EDUC_SEX_INTERACTION` | -1832.32    |
+| `cat__MARST`                | -5541.39    |
+| `cat__GQ`                   | -7956.89    |
+| `cat__SCHLTYPE`             | -16320.74   |
 :::
@@
 | Feature                  | VIF       |
-|-----------------------------|----------|
-| cat\_\_STATEICP             | 3.874599 |
-| cat\_\_GQ                   | 1.002446 |
-| cat\_\_MORTGAGE             | 2.614057 |
-| cat\_\_MARST                | 1.754080 |
-| cat\_\_SCHLTYPE             | 1.242432 |
-| cat\_\_OCC2010              | 2.953611 |
-| cat\_\_IND1990              | 6.566724 |
-| num\_\_AGE                  | 7.373021 |
-| num\_\_EDUC_SEX_INTERACTION | 7.066118 |
+|--------------------------|-----------|
+| cat__STATEICP            | 3.242254  |
+| cat__GQ                  | 1.002419  |
+| cat__MORTGAGE            | 2.658013  |
+| cat__MARST               | 1.741158  |
+| cat__SCHLTYPE            | 1.239412  |
+| cat__OCC2010             | 3.187731  |
+| cat__IND1990             | 8.490997  |
+| num__AGE                 | 7.871046  |
+| num__EDUC_SEX_INTERACTION| 6.945850  |
 :::
@@
 | Metric                     | Value              |
-|--------------------------------|------------------|
-| Mean Squared Error (MSE)       | 7,178,302,129.51 |
-| Root Mean Squared Error (RMSE) | 84,724.86        |
-| R-squared                      | 0.11099          |
-| Adjusted R-squared             | 0.11097          |
+|----------------------------|--------------------|
+| R-squared                  | 0.0774938018       |
+| Adjusted R-squared         | 0.0774637241       |
+| Mean Squared Error (MSE)   | 7,448,800,507.01   |
+| Root Mean Squared Error (RMSE) | 86,306.43        |
 :::

### 2024-10-25 paper/paper.qmd
Re-run the model & Updated model results with new data

@@
 |-----------------|------------------|
-| Best Model MSE  | 6,057,654,880.33 |
-| Best Model RMSE | 77,830.94        |
+| Best Model MSE  | 6,057,429,750.02 |
+| Best Model RMSE | 77,829.49        |
 :::
@@
 |-----------------|------------|
-| cat\_\_EDUC_new | 0.337465   |
-| cat\_\_IND1990  | 0.243701   |
-| num\_\_AGE      | 0.193615   |
-| cat\_\_SEX      | 0.149471   |
-| cat\_\_MARST    | 0.055754   |
-| cat\_\_MORTGAGE | 0.012897   |
-| cat\_\_SCHLTYPE | 0.004205   |
-| cat\_\_VETSTAT  | 0.002891   |
+| cat\_\_EDUC_new | 0.337065   |
+| cat\_\_IND1990  | 0.243454   |
+| num\_\_AGE      | 0.193926   |
+| cat\_\_SEX      | 0.149342   |
+| cat\_\_MARST    | 0.055948   |
+| cat\_\_MORTGAGE | 0.013064   |
+| cat\_\_SCHLTYPE | 0.004251   |
+| cat\_\_VETSTAT  | 0.002948   |
 | cat\_\_OWNERSHP | 0.000000   |
@@
 |--------------------------------|------------------|
-| Mean Squared Error (MSE)       | 5,901,467,894.81 |
-| Root Mean Squared Error (RMSE) | 76,821.01        |
+| Mean Squared Error (MSE)       | 5,191,183,577.52 |
+| Root Mean Squared Error (RMSE) | 72,049.87        |
 :::
@@
 |-----------------------|-----------|
-| Cross-Validation RMSE | 71,676.11 |
+| Cross-Validation RMSE | 71,994.55 |
 :::

### 2024-10-25 paper/paper.qmd
Change the position of title of tables in Model

@@
 ::: {#tbl-coefficients}
-**Table 1: Coefficients from the Linear Regression Model**
+\begin{center}
+Table 1: Coefficients from the Linear Regression Model
+\end{center}
 
@@
 ::: {#tbl-VIF}
-**Table 2: Variance Inflation Factors (VIF) for Predictor Variables**
+\begin{center}
+Table 2: Variance Inflation Factors (VIF) for Predictor Variables
+\end{center}
 
@@
 ::: {#tbl-modelperformance}
-**Table 3: Model Performance Metrics**
+\begin{center}
+Table 3: Model Performance Metrics
+\end{center}
 
@@
 ::: {#tbl-randomforestMSE}
-**Table 4: Best Random Forest Model Performance Metrics**
+\begin{center}
+Table 4: Best Random Forest Model Performance Metrics
+\end{center}
 
@@
 ::: {#tbl-randomforestimportance}
-**Table 5: Feature Importance in the Best Random Forest Model**
+\begin{center}
+Table 5: Feature Importance in the Best Random Forest Model
+\end{center}
 
@@
 ::: {#tbl-XGBoostMSE}
-**Table 6: XGBoost Model Performance Metrics**
+\begin{center}
+Table 6: XGBoost Model Performance Metrics
+\end{center}
 
@@
 ::: {#tbl-XGBoostCV}
-**Table 7: XGBoost Cross-Validation Performance**
+\begin{center}
+Table 7: XGBoost Cross-Validation Performance
+\end{center}
 

### 2024-10-25 paper/paper.qmd
Write the analysis in Results Section

@@
 |-----------------------------|-------------|
-| `cat__MORTGAGE`             | 11267.90    |
-| `num__AGE`                  | 415.91      |
-| `cat__STATEICP`             | 126.39      |
-| `cat__IND1990`              | -43.18      |
-| `cat__OCC2010`              | -139.55     |
-| `num__EDUC_SEX_INTERACTION` | -1832.32    |
-| `cat__MARST`                | -5541.39    |
-| `cat__GQ`                   | -7956.89    |
-| `cat__SCHLTYPE`             | -16320.74   |
+| cat__MORTGAGE             | 11267.90    |
+| num__AGE                  | 415.91      |
+| cat__STATEICP             | 126.39      |
+| cat__IND1990              | -43.18      |
+| cat__OCC2010              | -139.55     |
+| num__EDUC_SEX_INTERACTION | -1832.32    |
+| cat__MARST                | -5541.39    |
+| cat__GQ                   | -7956.89    |
+| cat__SCHLTYPE             | -16320.74   |
 :::
@@
 
-## Liner Regression Model
+## Linear Regression Model with Interaction Term
+
+### Linear Regression Results
+
+Table \ref{tbl-coefficients} provides the coefficients for each predictor in the model.
+The coefficients represent the estimated change in the target variable (income) for a one-unit change in the predictor, holding all other variables constant.
+
+The largest positive effect on income is observed for `cat_MORTGAGE`, which has a coefficient of 11,267.90.
+This suggests that individuals with a mortgage tend to have significantly higher incomes compared to those without, when other factors are held constant.
+Additionally, the coefficient for **`num__AGE`** is 415.91, indicating that income tends to increase with age, though the effect size is relatively small.
+
+On the other hand, some predictors have negative effects on income.
+The variable **`cat__SCHLTYPE`** (school type) has the largest negative coefficient of -16,320.74, suggesting that attending certain types of schools might be associated with lower income levels.
+Similarly, **`cat__GQ`** and **`cat__MARST`** (marital status) have negative coefficients of -7,956.89 and -5,541.39, respectively, implying that certain marital statuses or living arrangements are associated with lower income levels.
+
+Interestingly, the interaction term **`num__EDUC_SEX_INTERACTION`** has a negative coefficient of -1,832.32, suggesting that the relationship between education and income may differ by gender.
+The negative value could indicate that the combined effect of education and gender has a suppressing effect on income for certain groups.
+
+### Model Performance
+
+As shown in Table \ref{tbl-modelperformance}, the linear regression model explains a small portion of the variance in income, as evidenced by the R-squared value of 0.0775, which indicates that only 7.75% of the variability in income is explained by the predictors in the model.
+The Adjusted R-squared, which adjusts for the number of predictors, is slightly lower at 0.0775, confirming that the inclusion of additional predictors did not drastically improve the model’s explanatory power.
+
+Furthermore, the model’s Mean Squared Error (MSE) is 7,448,800,507, and the Root Mean Squared Error (RMSE) is 86,306.43, suggesting that the model's predictions deviate significantly from the actual income values.
+These performance metrics indicate that the linear regression model has limited predictive accuracy for this dataset.
+
+### Residual Analysis
+
+The residual analysis helps evaluate the linear regression model's performance.
+The residual histogram shows a strong right skew, indicating that while most predictions are close to the actual values, the model struggles with higher incomes.
+The long tail suggests significant under-prediction for some individuals.
+
+In the residuals vs.
+predicted values plot, the residuals exhibit a funnel shape, showing increasing variance as predicted income rises.
+This indicates heteroscedasticity, meaning the model's error increases for higher predicted values.
+Overall, the analysis suggests that the linear model fits well for lower incomes but underperforms for higher-income predictions, implying a need for more complex models to better capture these relationships.
+
+## Random Forest
+
+We employed a Random Forest model to predict income, tuning several hyperparameters to improve model performance.
+The best model was achieved with the following parameters: **max_depth=10**, **min_samples_leaf=4**, **min_samples_split=10**, and **n_estimators=300**.
+As we shown in Table \ref{tbl-randomforestMSE}, the final model had a Mean Squared Error (MSE) of 6,057,429,750 and a Root Mean Squared Error (RMSE) of 77,829.49, indicating a substantial improvement in prediction accuracy compared to the linear regression model.
+
+### Feature Importance
+
+The Random Forest model, Table \ref{tbl-randomforestimportance}, also allows for an examination of feature importance, which ranks the variables by their contribution to the model's predictions.
+The most important features in the model are:
+
+-   `cat__EDUC_new` (Education level) with an importance score of 0.337, indicating that education is the most significant factor affecting income in this model.
+
+-   `cat__IND1990` (Industry) has an importance score of 0.243, suggesting that the industry in which a person works is also a major determinant of income.
+
+-   `num__AGE` (Age) comes in third with a score of 0.194, implying that age plays a substantial role in income prediction.
+
+Other variables, such as `cat__SEX` (Sex) and `cat__MARST` (Marital Status), show moderate importance, while factors like `cat__MORTGAGE` and `cat__SCHLTYPE` (School Type) have much smaller impacts.
+Interestingly, `cat__OWNERSHP` (Home Ownership) has no contribution to the model, as its importance score is 0.000.
+
+### Residual Analysis
+
+The residuals from the Random Forest model were analyzed similarly to the linear model.
+The residuals plot, comparing predicted values with residuals, shows less heteroscedasticity compared to the linear model, suggesting that the Random Forest model handles the data more effectively, particularly for higher income levels.
+However, some degree of variability in the residuals still increases with predicted income, indicating that further tuning or model adjustments may be necessary for optimal performance.
+
+## XGBoost
+
+The XGBoost model, a gradient boosting algorithm, was optimized using a hyperparameter grid search, yielding the best parameters: **learning_rate=0.1**, **max_depth=6**, **min_child_weight=5**, and **n_estimators=300**.
+These parameters were selected to balance model complexity and overfitting control.
+As we shown in Table \ref{tbl-XGBoostMSE}, the final model achieved a Mean Squared Error (MSE) of 5,191,183,577.52 and a Root Mean Squared Error (RMSE) of 72,049.87, making it the best-performing model among the three evaluated.
+
+### Model Performance
+
+The XGBoost model’s Cross-Validation RMSE (Table \ref{tbl-XGBoostCV}) was 71,994.55, which aligns closely with the final RMSE, indicating that the model generalizes well across different data subsets.
+This consistency in performance suggests that XGBoost effectively captures the relationship between predictors and income without overfitting.
+
+### Residual Analysis
+
+The residuals vs.
+predicted values plot shows a similar pattern to the Random Forest model, with residuals clustering around zero for lower predicted values but spreading out for higher predicted incomes.
+However, compared to the Random Forest model, the XGBoost residuals show a slightly narrower spread, particularly at the higher income levels, indicating that XGBoost handles high-income predictions somewhat better.
+Nonetheless, some degree of heteroscedasticity remains, as the spread increases with predicted income.
 
@@
 
-# Appendix {#sec-appendix}
+# Appendix
 
-## Data Cleaning {#sec-data-cleaning}
+## Data Cleaning
 
-## Model Details {#sec-model-details}
+## Model Details
 

### 2024-10-25 paper/paper.qmd
Write Discussion Section

@@
 
-## Weaknesses and next steps
+In this analysis, we explored three different models—Linear Regression, Random Forest, and XGBoost—to predict income based on various demographic and categorical factors.
+Each model provided valuable insights into the relationship between income and predictors such as education, age, industry, and gender.
+Among the models, XGBoost demonstrated the best performance in terms of predictive accuracy, as evidenced by its lower RMSE and superior handling of complex data interactions.
+However, this does not mean it is without limitations or that further improvements cannot be made.
+
+## Key Findings
+
+One consistent finding across all models is the significant impact of education and industry on income.
+These factors were consistently ranked as the most important predictors in both Random Forest and XGBoost models.
+Age also played an important role but to a lesser extent, while variables such as marital status and sex showed moderate influence.
+
+The residual analysis revealed that none of the models perfectly fit the data, particularly for higher income levels.
+Both Random Forest and XGBoost exhibited a pattern of increasing residual variance (heteroscedasticity) as predicted income increased, which suggests that the models struggled to accurately predict high-income values.
+Linear Regression, while more interpretable, performed the worst in terms of accuracy, especially for non-linear relationships between the predictors and income.
+
+Despite the overall success of XGBoost, it is clear that there are areas where the model's performance could be further refined.
+
+## Weaknesses and Limitations
+
+One major limitation encountered throughout the analysis was the skewness of the income distribution.
+The income data is heavily left-skewed, with most individuals earning relatively low amounts and a few individuals earning substantially higher incomes.
+This skewness likely contributed to the residual patterns we observed, particularly in the linear and Random Forest models.
+
+We attempted to address this by applying a logarithmic transformation to income in an effort to normalize the data.
+However, this transformation did not lead to substantial improvements in model performance.
+Even after the transformation, the residual patterns remained, indicating that the underlying relationships between predictors and income might require more sophisticated modeling approaches than simple transformations can provide.
+
+Another potential solution—though not ideal—would be to exclude extreme income values from the analysis.
+Removing high-income outliers might reduce the heteroscedasticity issue and result in a better fit for the majority of the population.
+However, excluding high-income individuals would also mean losing valuable information about the upper end of the income distribution, which is crucial for a complete analysis.
+Before taking this step, it would be necessary to justify it based on domain-specific knowledge or further data exploration.
+
+## Next Steps
+
+Looking forward, there are several potential avenues for further research and model improvement:
+
+-   While we focused on standard predictors like age, education, and industry, additional variables or interaction terms might improve model performance.
+For example, interactions between age and education, or industry and gender, could capture more nuanced relationships.
+
+-    Instead of removing high-income individuals, more sophisticated outlier treatment techniques could be applied.
+For instance, applying quantile regression or using robust regression models might help better model extreme values without distorting the broader data patterns.
+
+-   The skewness in income distribution poses a significant challenge for modeling.
+Exploring techniques like synthetic data generation for high-income individuals or applying weighting schemes to account for data imbalance might help the models better capture the variance across all income levels.
+
+In conclusion, while XGBoost provided the most accurate results, there is room for improvement in handling skewed data and better modeling high-income individuals.
+Future work will focus on addressing these issues while exploring more advanced modeling techniques and feature engineering to enhance predictive performance further.
 

### 2024-10-26 paper/paper.qmd
Add graphs into paper and perform cross-ref, update the name of XGBoost Residual Plot

@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{../other/graphs/LinearModelResidual1.png}
+\caption{Linear Model Residuals Plot}
+\label{fig:residual-linear1}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{../other/graphs/LinearModelResidual2.png}
+\caption{Linear Model Residuals vs. Predicted Values Plot}
+\label{fig:residual-linear2}
+\end{figure}
+
 We use the statsmodels package [@cite_statsmodels] to calculate the VIF for each predictor.
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{../other/graphs/RandomForestResidual.png}
+\caption{Random Forest Model Residuals vs. Predicted Values Plot}
+\label{fig:residual-randomforest}
+\end{figure}
+
 The feature importances were extracted from the best model using RandomForestRegressor from scikit-learn [@cite_sklearn] and displayed using pandas [@pandas].
@@
 We explored a range of hyperparameters to find the optimal configuration for our dataset.
-The grid search included the learning rate (`learning_rata`), max depth (`max_depth`), number of trees (`n_estimators)`, and minimum child weight (`min_child_weight`).
+The grid search included the learning rate (`learning_rata`), max depth (`max_depth`), number of trees (`n_estimators`), and minimum child weight (`min_child_weight`).
 Specifically, we tested learning rates of 0.01, 0.05, and 0.1, tree depths of 4, 6, and 8, and `n_estimators` values of 100, 200, and 300.
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{../other/graphs/XGBoostResidual.png}
+\caption{XGBoost Model Residuals vs. Predicted Values Plot}
+\label{fig:residual-XGBoost}
+\end{figure}
+
 We also performed cross-validation to ensure the robustness of the XGBoost model.
@@
 This suggests that individuals with a mortgage tend to have significantly higher incomes compared to those without, when other factors are held constant.
-Additionally, the coefficient for **`num__AGE`** is 415.91, indicating that income tends to increase with age, though the effect size is relatively small.
+Additionally, the coefficient for `num__AGE` is 415.91, indicating that income tends to increase with age, though the effect size is relatively small.
 
 On the other hand, some predictors have negative effects on income.
-The variable **`cat__SCHLTYPE`** (school type) has the largest negative coefficient of -16,320.74, suggesting that attending certain types of schools might be associated with lower income levels.
-Similarly, **`cat__GQ`** and **`cat__MARST`** (marital status) have negative coefficients of -7,956.89 and -5,541.39, respectively, implying that certain marital statuses or living arrangements are associated with lower income levels.
+The variable `cat__SCHLTYPE` (school type) has the largest negative coefficient of -16,320.74, suggesting that attending certain types of schools might be associated with lower income levels.
+Similarly, `cat__GQ` and `cat__MARST` (marital status) have negative coefficients of -7,956.89 and -5,541.39, respectively, implying that certain marital statuses or living arrangements are associated with lower income levels.
 
-Interestingly, the interaction term **`num__EDUC_SEX_INTERACTION`** has a negative coefficient of -1,832.32, suggesting that the relationship between education and income may differ by gender.
+Interestingly, the interaction term `num__EDUC_SEX_INTERACTION` has a negative coefficient of -1,832.32, suggesting that the relationship between education and income may differ by gender.
 The negative value could indicate that the combined effect of education and gender has a suppressing effect on income for certain groups.
@@
 The residual analysis helps evaluate the linear regression model's performance.
-The residual histogram shows a strong right skew, indicating that while most predictions are close to the actual values, the model struggles with higher incomes.
+The residual histogram as shown in Figure \ref{fig:residual-linear1} shows a strong right skew, indicating that while most predictions are close to the actual values, the model struggles with higher incomes.
 The long tail suggests significant under-prediction for some individuals.
@@
 In the residuals vs.
-predicted values plot, the residuals exhibit a funnel shape, showing increasing variance as predicted income rises.
+predicted values plot as shown in Figure \ref{fig:residual-linear2}, the residuals exhibit a funnel shape, showing increasing variance as predicted income rises.
 This indicates heteroscedasticity, meaning the model's error increases for higher predicted values.
@@
 
-The residuals from the Random Forest model were analyzed similarly to the linear model.
+The residuals from the Random Forest model as shown in Figure \ref{fig:residual-randomforest} were analyzed similarly to the linear model.
 The residuals plot, comparing predicted values with residuals, shows less heteroscedasticity compared to the linear model, suggesting that the Random Forest model handles the data more effectively, particularly for higher income levels.
@@
 The residuals vs.
-predicted values plot shows a similar pattern to the Random Forest model, with residuals clustering around zero for lower predicted values but spreading out for higher predicted incomes.
+predicted values plot as shown in Figure \ref{fig:residual-XGBoost} shows a similar pattern to the Random Forest model, with residuals clustering around zero for lower predicted values but spreading out for higher predicted incomes.
 However, compared to the Random Forest model, the XGBoost residuals show a slightly narrower spread, particularly at the higher income levels, indicating that XGBoost handles high-income predictions somewhat better.

### 2024-10-26 paper/paper.qmd
Remove the model details section
Finish Appendix Data Cleaning Section

@@
 We implemented three models: Linear Regression with Interaction Terms, Random Forest, and XGBoost, each chosen for their unique strengths in prediction and interpretation.
-
-Background details are included in Appendix @sec-model-details.
 
@@
 
-## Data Cleaning
+## Data Cleaning {#sec-datacleaning}
 
-## Model Details
+In the data cleaning process, we first examined all required variables and removed any observations with missing values (N/A).
+Notably, the variable `CITY`, which we initially intended to include in the analysis, was excluded due to its high proportion of missing values, which compromised data completeness.
+
+For the response variable, personal total income (INCTOT), we identified several non-informative coded values, such as `0000000 = None`, `9999998 = Unknown`, and `-009995 = -$9,900 (1980)` and so on.
+These values lack meaningful information for analysis, so we removed any records containing such entries to maintain the integrity of the income data.
+
+In terms of predictor variables, we found that several included codes like "not applicable" or "does not exist," which represent cases without relevant data for our analysis.
+These values were also removed to ensure only relevant data points were used in modeling.
+
+Furthermore, we simplified the `EDUC` variable (education level), which originally included highly granular categories (e.g., fifth grade, sixth grade, second year of college, fifth year of college).
+Such detailed classifications were unnecessary for our purposes, so we re-coded `EDUC` into a new variable, `EDUC_new`, with broader categories:
+
+-   'No Schooling': 0,
+-   'Nursery School': 1,
+-   'Primary School': 2,
+-   'Middle School': 3,
+-   'College 1-4 Years': 4,
+-   'College 5+ Years': 5
+
+This simplified categorization enhances interpretability and usability in the analysis.
+By conducting these data cleaning steps, we produced a refined and consistent dataset, ready for accurate and reliable modeling.
 

### 2024-10-26 paper/paper.qmd
Deleted variable "OWNERSHP".

@@
 
-The dataset used here is downloaded from the Toronto Open Data website via @TorontoOpenData.
-This dataset contains all the reported crimes that happened in Toronto from 2014 to 2023.
-This dataset is grouped by the year of the reported crime, its category and belonging subtype, and the count of the subtype being reported and cleared for that year for each division.
-Because I am examining the crime pattern in the city, I have dropped the division information and aggregated the existing data according to their subtype and the year of the crime being reported.
-In the following subsections, I will review all the variables used in this report and provide some basic descriptive statistics.
-The first five rows of the cleaned data used for analysis are attached (@tbl-table1).
+The dataset used here is downloaded from the IPUMS USA website via @ipums_usa_v15.
+Our project's extracted dataset contains all the responses from the American Community Survey (ACS) from 2023.
+Our extraction included the following variables: total income, dwelling mortgage status, sex of the individual, marital status, education attainment, private or public school, industry or occupation, veteran status, and age.
+We acknowledge that some of the variables (such as educational level, occupation, etc.) we have chosen here may have a more detailed version that contains more information.
+However, due to the nature of categorical variables, having detailed variables will create more classes of that variable, thus leading to more features for our model when encoding those variables.
+Although more features will improve the overall performance of our predictive model, it will also create issues such as overfitting, needing more computational resources, and lack of interpretability.
+Due to the trade-off of these mentioned pros and cons, we have decided to proceed with these higher-level variables for this project for simplicity.
+Furthermore, we have cleaned the data for better data quality.
+Details of our data cleaning processes are provided in @sec-datacleaning.
+In the following subsections, we will review all the variables used in this report and provide some basic descriptive statistics.
 
-```{python}
-#| label: tbl-table1
-#| tbl-cap: Example of Cleaned Data
-#| warning: false
-#| echo: false
-
-import pandas as pd
-data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
-data.index = data.index + 1
-data.head()
-# styled_data = data.head().style.format({'ClearRate': "{:.2f}"}).hide().set_table_styles([{"selector":"th.col_heading,td","props":[("width","100px")]}])
-# styled_data
-#data.values[:5]
-#print(data.head().to_string(index=False))
-
-```
-
-## Report Year
-The report year variable is the number of crimes being reported.
-In this dataset, the data spans from 2014 to 2023, encompassing ten years.
-No month or date information was given; thus, there are only ten different values for this variable in chronological order.
-
-## Category
-The category includes information about the nature of the crime.
-There are six crime categories: Crimes Against Property, Crimes Against the Person, Other Federal Statute Violations, Other Criminal Code Violations, Controlled Drugs and Substances Act, and Criminal Code Traffic.
-They are listed in the table below (@tbl-table2).
+This is a general version (compared to the detailed version). The reason of choosing the general version is mentioned in @sec-data above
+## Mortgage
+The Mortgage variable describes the mortgage status of dwelling.
+There are five different values, they are "N/A", "No, owned free and clear", "Check mark on manuscript (probably yes)", "Yes, mortgaged/ deed of trust or similar debt", and "Yes, contract to purchase".
+We have dropped the N/A values for better data quality.
+Details available in @sec-datacleaning.
+The count of the remaining classes is shown below in (@fig-figure1).
 
@@
 #| echo: false
-#| label: tbl-table2
+#| label: fig-figure1
 #| warning: false
-#| tbl-cap: Six Crime Categories
-# import pandas as pd
-# import plotly.graph_objects as go
-# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
-# unique_category = data.Category.unique()
-# category_df = pd.DataFrame(unique_category, columns=['Categories'])
-# category_df.index = category_df.index + 1
-# category_df
+#| tbl-cap: Dwelling Ownership Details
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+plt.figure(figsize=(10, 6))
+sns.countplot(x='MORTGAGE', data=data, palette='pastel', edgecolor='black')
+plt.title('Distribution of Mortgage Status', fontsize=14)
+plt.xlabel('Mortgage Status', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(rotation=30, ha='right')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
 ```
 
-## Subtype
+## Sex
 There exist multiple subtypes under each crime category. The following is an exhaustive table (@tbl-table3) of all crimes' subtypes and their respective category.
@@
 
-## Count
+## Age
 In the original table, this value is grouped by the subtype of the crime, the division, and the year when the crime was reported.
@@
 
-## Count_Cleared
+## Marrital Status
 These are the counts of crimes identified as cleared.
@@
 
-## Case_Clearing_Rate
+## Educational Attainment
+This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
+A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
+The value is ranged from 0 to 100%.
+
+## School Type
+This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
+A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
+The value is ranged from 0 to 100%.
+
+## Industry
+This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
+A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
+The value is ranged from 0 to 100%.
+
+## Vateran Status
+This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
+A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
+The value is ranged from 0 to 100%.
+
+## Total Income
 This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.

### 2024-10-26 paper/paper.qmd
Review the result with updated data

@@
 |-----------------|------------------|
-| Best Model MSE  | 6,057,429,750.02 |
-| Best Model RMSE | 77,829.49        |
+| Best Model MSE  | 6,057,426,411.65 |
+| Best Model RMSE | 77,829.47        |
 :::
@@
 | cat\_\_VETSTAT  | 0.002948   |
-| cat\_\_OWNERSHP | 0.000000   |
 :::
@@
 |--------------------------------|------------------|
-| Mean Squared Error (MSE)       | 5,191,183,577.52 |
-| Root Mean Squared Error (RMSE) | 72,049.87        |
+| Mean Squared Error (MSE)       | 5,854,629,848.13 |
+| Root Mean Squared Error (RMSE) | 76,515.55        |
 :::
@@
 \begin{center}
-Table 7: XGBoost Cross-Validation Performance
+Table 7: Cross-Validation RMSE for XGBoost Model
 \end{center}
 
-| Metric                | Value     |
-|-----------------------|-----------|
-| Cross-Validation RMSE | 71,994.55 |
+| Fold  | RMSE       |
+|-------|------------|
+| 1     | 76,693.86  |
+| 2     | 75,692.41  |
+| 3     | 76,193.42  |
+| 4     | 77,365.82  |
+| 5     | 76,552.10  |
+| **Average** | **76,499.52** |
 :::
@@
 The best model was achieved with the following parameters: **max_depth=10**, **min_samples_leaf=4**, **min_samples_split=10**, and **n_estimators=300**.
-As we shown in Table \ref{tbl-randomforestMSE}, the final model had a Mean Squared Error (MSE) of 6,057,429,750 and a Root Mean Squared Error (RMSE) of 77,829.49, indicating a substantial improvement in prediction accuracy compared to the linear regression model.
+As we shown in Table \ref{tbl-randomforestMSE}, the final model had a Mean Squared Error (MSE) of 6,057,426,411.65 and a Root Mean Squared Error (RMSE) of 77,829.47, indicating a substantial improvement in prediction accuracy compared to the linear regression model.
 
@@
 Other variables, such as `cat__SEX` (Sex) and `cat__MARST` (Marital Status), show moderate importance, while factors like `cat__MORTGAGE` and `cat__SCHLTYPE` (School Type) have much smaller impacts.
-Interestingly, `cat__OWNERSHP` (Home Ownership) has no contribution to the model, as its importance score is 0.000.
 
@@
 These parameters were selected to balance model complexity and overfitting control.
-As we shown in Table \ref{tbl-XGBoostMSE}, the final model achieved a Mean Squared Error (MSE) of 5,191,183,577.52 and a Root Mean Squared Error (RMSE) of 72,049.87, making it the best-performing model among the three evaluated.
+As we shown in Table \ref{tbl-XGBoostMSE}, the final model achieved a Mean Squared Error (MSE) of 5,854,629,848.13 and a Root Mean Squared Error (RMSE) of 76,515.55, making it the best-performing model among the three evaluated.
 
@@
 
-The XGBoost model’s Cross-Validation RMSE (Table \ref{tbl-XGBoostCV}) was 71,994.55, which aligns closely with the final RMSE, indicating that the model generalizes well across different data subsets.
+The XGBoost model’s Cross-Validation RMSE (Table \ref{tbl-XGBoostCV}) was 76,499.52, which aligns closely with the final RMSE, indicating that the model generalizes well across different data subsets.
 This consistency in performance suggests that XGBoost effectively captures the relationship between predictors and income without overfitting.

### 2024-10-26 paper/paper.qmd
Data Part Complete

@@
 
-This is a general version (compared to the detailed version). The reason of choosing the general version is mentioned in @sec-data above
 ## Mortgage
@@
 #| warning: false
-#| tbl-cap: Dwelling Ownership Details
+#| fig-cap: Dwelling Ownership Details
 
@@
 import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
 plt.figure(figsize=(10, 6))
 sns.countplot(x='MORTGAGE', data=data, palette='pastel', edgecolor='black')
-plt.title('Distribution of Mortgage Status', fontsize=14)
 plt.xlabel('Mortgage Status', fontsize=12)
@@
 ## Sex
-There exist multiple subtypes under each crime category. The following is an exhaustive table (@tbl-table3) of all crimes' subtypes and their respective category.
+This variable is coded as a binary variable, and we will use it as-is in our project.
+This does not suggest that the authors agree with this type of sex classification, but rather that it aligns with the available data and serves the research purposes of this project.
+The binary coding is a practical choice to maintain consistency with the dataset, though it may not fully capture the complexity and diversity of gender identities.
+@fig-figure2 below shows the distribution of this variable.
 
 ```{python}
+
 #| echo: false
-#| label: tbl-table3
+#| label: fig-figure2
 #| warning: false
-#| tbl-cap: Crime Categories and Subtypes
-# import pandas as pd
-# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
-# unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
-# unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
-# unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
-# unique_category_subtype_df.reset_index(drop=True, inplace=True)
-# unique_category_subtype_df.index = unique_category_subtype_df.index + 1
-# unique_category_subtype_df
+#| fig-cap: Sex Distribution
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+plt.figure(figsize=(10, 6))
+sns.countplot(x='SEX', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Sex', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(ha='right')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
 ```
@@
 ## Age
-In the original table, this value is grouped by the subtype of the crime, the division, and the year when the crime was reported.
-The original count indicates the number of a specific subtype of crime reported within a particular division for the year.
-However, as mentioned before, because I am only interested in all the crimes in the City of Toronto, I have dropped the division information and aggregated the count from all the divisions to a single value.
-Therefore, for each subtype of the crimes, a total count of that subtype is reported in a single year.
+Age describes the age of the individual, and a age of 0 was coded for the individuals under 1 year old.
+For our dataset, the minimum age is 16, however.
+Age distribution of our dataset is shown below in @fig-figure3.
+```{python}
 
-## Marrital Status
-These are the counts of crimes identified as cleared.
-In plain words, these are crimes that are dealt/solved.
-I have taken the same approach for this column as the previous one.
-After cleaning the data,f, or each subtype of the crimes, there is a total count of that subtype being reported that is also cleared in a single year.
+#| echo: false
+#| label: fig-figure3
+#| warning: false
+#| fig-cap: Age Distribution
 
-## Educational Attainment
-This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
-A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
-The value is ranged from 0 to 100%.
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+plt.figure(figsize=(10, 6))
+sns.countplot(x='AGE', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Age', fontsize=12)
+plt.ylabel('Count', fontsize=8)
+plt.xticks(rotation=90, ha='center')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
+```
+
+## Marital Status
+Marital status describes the marital status of the individual.
+There are 7 different values, namely "Married, spouse present", "Married, spouse absent", "Separated", "Divorced", "Widowed", "Never married/single", and "Blank, missing".
+We have dropped the N/A values for better data quality.
+Details available in @sec-datacleaning.
+The following @fig-figure4 illustrate the distribution of marital status of the participants in our dataset.
+```{python}
+
+#| echo: false
+#| label: fig-figure4
+#| warning: false
+#| fig-cap: Counts of Marital Status
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+marrital_status_label = {
+  '1': 'Married, spouse present',
+  '2': 'Married, spouse absent',
+  '3': 'Separated',
+  '4': 'Divorced',
+  '5': 'Widowed',
+  '6': 'Never married/single',
+  '9': 'Blank, missing'
+}
+data['MARST'] = data['MARST'].map(marrital_status_label)
+plt.figure(figsize=(10, 6))
+sns.countplot(x='MARST', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Marital Status', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(rotation=30, ha='center')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
+```
+
+## Educational Attainment [general version]
+This is a general version (compared to the detailed version).
+The reason of choosing the general version is mentioned in @sec-data above.
+This column represents the highest educational attainment achieved by the individual.
+The distribution of the variable is shown in @fig-figure5.
+```{python}
+
+#| echo: false
+#| label: fig-figure5
+#| warning: false
+#| fig-cap: Educational Attainment
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+educ_new_label = {
+    0: 'No Schooling',
+    1: 'Nursery School',
+    2: 'Primary School',
+    3: 'Middle School',
+    4: 'College 1-4 Years',
+    5: 'College 5+ Years'
+}
+data['EDUC_new'] = data['EDUC_new'].map(educ_new_label)
+plt.figure(figsize=(10, 6))
+sns.countplot(x='EDUC_new', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Educational Attainment', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(rotation=30, ha='center')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
+```
 
 ## School Type
-This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
-A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
-The value is ranged from 0 to 100%.
+School type indicates if the individual went to a public school or private school.
+Other than missing values, there are three different classes.
+They are "Not enrolled", "Public school", and "Private school".
+We are not sure how the participants would respond if they have been to more than one type of schools.
+@fig-figure6 shows the distribution of our sample.
+```{python}
 
+#| echo: false
+#| label: fig-figure6
+#| warning: false
+#| fig-cap: School Type
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+school_type_label = {
+    "1": 'Not Enrolled',
+    "2": 'Private School',
+    "3": 'Public School',
+}
+data['SCHLTYPE'] = data['SCHLTYPE'].map(school_type_label)
+plt.figure(figsize=(10, 6))
+sns.countplot(x='SCHLTYPE', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Private or Public School', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(rotation=30, ha='center')
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
+```
 ## Industry
-This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
-A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
-The value is ranged from 0 to 100%.
+Industry describes the type of industry the individual works in.
+Same as the educational attainment variable, more detailed version of this variable is available (i.e., Occupation).
+There are more than 200 industries listed in our sample.
+A detailed description can be seen at [IPUMS USA](https://usa.ipums.org/usa-action/variables/IND1990#description_section).
 
-## Vateran Status
-This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
-A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
-The value is ranged from 0 to 100%.
+## Veteran Status [General Version]
+Veteran status idicates whether the individual is a veteran or not.
+It can be either "Yes", "No", or "N/A".
+Please notice that "N/A" here is different from missing, suggesting this might be a different indicator, which is not explicitly explained from the ddl file provided.
+There is also a detailed version of this variable, which is not used in our project.
+The counts is displayed in @tbl-table1.
+```{python}
+
+#| echo: false
+#| label: tbl-table1
+#| warning: false
+#| fig-cap: Veteran Status
+
+import pandas as pd
+import matplotlib.pyplot as plt
+
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+vet_status_labels = {'0': 'N/A', '1': 'No', '2': 'Yes'}
+data['VETSTAT'] = data['VETSTAT'].map(vet_status_labels)
+vet_status_count = data['VETSTAT'].value_counts().reset_index()
+vet_status_count.columns = ['Veteran Status', 'Count']
+fig, ax = plt.subplots(figsize=(6, 2))
+ax.axis('off')
+table = ax.table(
+    cellText=vet_status_count.values,
+    colLabels=vet_status_count.columns,
+    cellLoc='center',
+    loc='center'
+)
+table.auto_set_font_size(False)
+table.set_fontsize(10)
+table.scale(1, 1.5)
+plt.tight_layout()
+plt.show()
+```
 
 ## Total Income
-This column was not included in the raw dataset but was created by dividing the cleared crimes by total crimes.
-A higher case-clearing rate for a particular subtype of crime usually suggests a higher effectiveness of law enforcement in dealing with this subtype of crime.
-The value is ranged from 0 to 100%.
+Total income is the total income made by the individual for the year.
+The distribution of total income is depicted in @fig-figure7.
+We have removed the outliers from the graph to have a better understanding of the general pattern of the overall distribution.
+```{python}
 
+#| echo: false
+#| label: fig-figure7
+#| warning: false
+#| fig-cap: Total Income Distribution
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+plt.figure(figsize=(10, 6))
+sns.histplot(x='INCTOT', data=data, palette='pastel', edgecolor='black')
+plt.xlabel('Total Income', fontsize=12)
+plt.ylabel('Count', fontsize=12)
+plt.xticks(rotation=30, ha='center')
+plt.xlim(0, 250000)
+plt.grid(axis='y', linestyle='--', alpha=0.7)
+plt.tight_layout()
+plt.show()
+
+```
 # Exploratory Data Analysis {#sec-eda}
@@
 \begin{center}
-Table 1: Coefficients from the Linear Regression Model
+Table 2: Coefficients from the Linear Regression Model
 \end{center}
@@
 
-*Table 1: Linear Regression Coefficients Table.*
+*Table 2: Linear Regression Coefficients Table.*
 
@@
 \begin{center}
-Table 2: Variance Inflation Factors (VIF) for Predictor Variables
+Table 3: Variance Inflation Factors (VIF) for Predictor Variables
 \end{center}
@@
 
-*Table 2: This table displays the Variance Inflation Factors (VIF) for each of the predictor variables in the model. VIF values greater than 10 indicate potential multicollinearity issues.*
+*Table 3: This table displays the Variance Inflation Factors (VIF) for each of the predictor variables in the model. VIF values greater than 10 indicate potential multicollinearity issues.*
 
@@
 \begin{center}
-Table 3: Model Performance Metrics
+Table 4: Model Performance Metrics
 \end{center}
@@
 
-*Table 3: This table presents the performance metrics of the linear regression model, including MSE, RMSE, R-squared, and Adjusted R-squared.
+*Table 4: This table presents the performance metrics of the linear regression model, including MSE, RMSE, R-squared, and Adjusted R-squared.
 These metrics are used to assess the model's fit and predictive accuracy.*
@@
 \begin{center}
-Table 4: Best Random Forest Model Performance Metrics
+Table 5: Best Random Forest Model Performance Metrics
 \end{center}
@@
 
-*Table 4: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the best-tuned Random Forest model, demonstrating the model's performance after hyperparameter tuning.*
+*Table 5: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the best-tuned Random Forest model, demonstrating the model's performance after hyperparameter tuning.*
 
@@
 \begin{center}
-Table 5: Feature Importance in the Best Random Forest Model
+Table 6: Feature Importance in the Best Random Forest Model
 \end{center}
@@
 
-*Table 5: This table lists the feature importance values for the best-performing Random Forest model.
+*Table 6: This table lists the feature importance values for the best-performing Random Forest model.
 The higher the importance value, the more significant the feature is in predicting the target variable (income).*
@@
 \begin{center}
-Table 6: XGBoost Model Performance Metrics
+Table 7: XGBoost Model Performance Metrics
 \end{center}
@@
 
-*Table 6: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the final XGBoost model, reflecting its predictive accuracy.*
+*Table 7: This table shows the Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) for the final XGBoost model, reflecting its predictive accuracy.*
 
@@
 \begin{center}
-Table 7: Cross-Validation RMSE for XGBoost Model
+Table 8: XGBoost Cross-Validation Performance
 \end{center}
@@
 
-*Table 7: This table shows the Root Mean Squared Error (RMSE) from cross-validation, demonstrating the model’s performance across different subsets of the data.*
+*Table 8: This table shows the Root Mean Squared Error (RMSE) from cross-validation, demonstrating the model’s performance across different subsets of the data.*
 

### 2024-10-26 paper/paper.qmd
Finish EDA in paper and revise the graph position in Data Section

@@
 #| warning: false
-#| fig-cap: Dwelling Ownership Details
+#| message: false
+#| fig-pos: "H"
+#| fig-cap: Mortgage Status Distribution
 
@@
 data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
-plt.figure(figsize=(10, 6))
+plt.figure(figsize=(5, 3))
 sns.countplot(x='MORTGAGE', data=data, palette='pastel', edgecolor='black')
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Sex Distribution
@@
 data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
-plt.figure(figsize=(10, 6))
+plt.figure(figsize=(5, 3))
 sns.countplot(x='SEX', data=data, palette='pastel', edgecolor='black')
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Age Distribution
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Counts of Marital Status
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Educational Attainment
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: School Type
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Veteran Status
@@
 #| warning: false
+#| message: false
+#| fig-pos: "H"
 #| fig-cap: Total Income Distribution
@@
 # Exploratory Data Analysis {#sec-eda}
+
+## Income Distribution by Mortgage Status
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../other/graphs/EDAincomebymortgage.png}
+    \caption{Income Distribution by Mortgage Status}
+    \label{fig:income-mortgage}
+\end{figure}
+
+Figure \ref{fig:income-mortgage} illustrates the income distribution for individuals with different mortgage statuses.
+The highest average income is observed among individuals with a "Mortgage/Trust Deed," followed by those who own their homes outright ("Fully Owned").
+Those with a "Signed Purchase Contract" have the lowest average income among the groups.
+This pattern may reflect a correlation between mortgage status and income level, where individuals with higher incomes are more likely to secure traditional mortgage loans, while those with lower incomes tend towards alternative purchasing arrangements.
+This could indicate the influence of financial stability and access to credit on mortgage status.
+
+## Correlation Analysis of Age, Education, and Income
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../other/graphs/EDAcorrelation.png}
+    \caption{Correlation Matrix of Age, Education, and Income}
+    \label{fig:correlation-matrix}
+\end{figure}
+
+Figure \ref{fig:correlation-matrix} shows the correlation matrix for three key variables: age, education level (EDUC_new), and total personal income (INCTOT).
+The matrix reveals that there is a moderate positive correlation between education level and income (0.29), suggesting that higher education levels are generally associated with higher income.
+Additionally, a smaller positive correlation exists between age and income (0.14), which might indicate that as individuals grow older, they tend to earn more due to accumulated experience.
+The correlation between age and education (0.06) is minimal, which is expected as these two variables are less directly related.
+Overall, this analysis supports the notion that education level plays a more significant role in income variation than age within this dataset.
+
+## Average Income by Education and Gender {#sec-eda3}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../other/graphs/EDAinteraction.png}
+    \caption{Average Income by Education and Sex}
+    \label{fig:income-education-sex}
+\end{figure}
+
+Figure \ref{fig:income-education-sex} presents the average total personal income based on different levels of education for both male and female groups.
+From the plot, it is observed that individuals with higher levels of education, particularly those who attended "College 5+ Years," tend to earn more on average compared to other groups.
+However, there are some variations in income within each education level, as indicated by the error bars.
+Generally, males tend to have higher average incomes across most education levels.
+This difference suggests a possible gender-based income disparity, even after controlling for education levels.
+Additionally, the variability in income among education levels highlights that factors beyond educational attainment, such as occupation and industry, may further impact income.
+Thus, incorporating an interaction term between education and gender in the model is necessary to capture these nuanced effects more accurately.
 
@@
 
-Note that according to our previous analysis in @sec-eda, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
+Note that according to our previous analysis in @sec-eda3, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
 This allowed us to model the non-linear relationship between these two variables and their joint impact on the outcome variable.

### 2024-11-18 paper/paper.qmd
Add alpha on the residual plots and modify the paper

@@
 ---
-title: "Personal Total Income of Individual in the United States of America in 2023"
-subtitle: "A Predictive Model Using the American Community Survey Data"
+title: "Predicting Personal Total Income of Individuals in the United States in 2023"
+subtitle: "XGBoost Captures the Complex Pattern Better Than Random Forest and Linear Regression"
 author: 
   - Justin (Jiazhou) Bi and Weiyang Li
-thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/ACS-Income-Model/tree/main."
-date: Oct 25, 2024
+thanks: "All project related files available at: https://github.com/Jiazhou-Bi/ACS-Income-Model"
+date: Nov 14, 2024
 date-format: long
@@
 Income is a fundamental concept as it affects nearly every member of society, influencing access to resources, quality of life, and overall economic stability.
-At the same time, it is also a highly complex concept to define, as it can be interpreted differently depending on various perspectives, such as economic, social, legal, and tax purposes.
+At the same time, it is also a difficult concept to define, as it can be interpreted differently depending on various perspectives, such as economic, social, legal, and tax purposes.
 Therefore, we must examine various demographical factors that may impact income to understand its concept better.
-Our project defines income as "all forms of net financial resources generated or lost from work, investments, or other activities." This project aims to predict personal income using data from the 2023 American Community Survey (ACS) (@ipums_usa_v15).
-The ACS provides detailed demographic information, including age, education, employment, marital status, mortgage status, veteran status, etc...
+Our project defines income as "all forms of net financial resources generated or lost from work, investments, or other activities." This project aims to predict personal income using data from the 2023 American Community Survey (ACS) [@ipums_usa_v15].
+The ACS provides detailed demographic information, including age, education, employment, marital status, mortgage status, veteran status.
 As a result, it provides a robust dataset for studying potential factors that may influence income.
@@
 As such, the results may vary across different periods, and their generalizability has not been assessed within this project.
-The raw data used in this analysis was sourced from IPUMS using the IPUMS API.
-(@ipums_api) We utilized pandas (@pandas) for data processing, and data visualization was carried out using Seaborn.
-(@seaborn) Additionally, all tables presented in this report were generated with Plotly.
-(@plotly)
+The raw data used in this analysis was sourced from IPUMS using the IPUMS API [@ipums_api].
+We utilized pandas [@pandas] for data processing, and data visualization was carried out using Seaborn [@seaborn].
+Additionally, all tables presented in this report were generated with Plotly [@plotly].
 
 Income inequality is a serious problem and can lead to significant consequences and it may affect social structure, economic stability, as well as policy-making processes.
-Researchers found that demographic factors, such as education level and occupation type, can significantly impact an individual's financial standing.
-(@Autor_2013) Similarly, social relationships, such as marital status and household characteristics, affect income dynamics.
-(@WaiteGallagher_2000) Therefore, applying machine learning models to analyze the ACS dataset and using the information mentioned to predict personal income is reasonable.
+Researchers found that demographic factors, such as education level and occupation type, can significantly impact an individual's financial standing [@Autor_2013].
+Similarly, social relationships, such as marital status and household characteristics, affect income dynamics [@WaiteGallagher_2000].
+Therefore, applying machine learning models to analyze the ACS dataset and using the information mentioned to predict personal income is reasonable.
 
-We have chosen 2023 ACS dataseet because it offers many demographical details, as well as self-reported total income of the individuals.
+We have chosen 2023 ACS data because it offers many demographical details, as well as self-reported total income of the individuals.
 The sample is also representative and covers all geographical locations in the USA.
 Furthermore, it provides easy access to their database through API.
-We will employ advanced predictive models, including random forest, lineaer regression, and extreme gradient boosting machine to estimate personal income based on demographic factors to provide insights into which potential factors have the most significant impact on personal income.
+We use predictive models, including random forest, linear regression, and extreme gradient boosting machine to estimate personal income based on demographic factors to provide insights into which potential factors have the most significant impact on personal income.
 We hope the findings generated from our project can provide insights to policy-making process so that we can, hopefully, reduce income inequality and support further economic development in the future.
@@
 
+In this analysis, we explored three different models—Linear Regression, Random Forest, and XGBoost—to predict income based on various demographic factors.
+Each model provides valuable insights into the relationships between income and predictors, such as education, age, industry, and gender.
+Among the models, XGBoost demonstrated the best performance in terms of predictive accuracy, as evidenced by its lower RMSE and superior handling of complex data interactions.
+
+The remainder of this paper is structured as follows: @sec-data introduces the raw dataset and describes the cleaned datasets, along with a preliminary analysis through numerical summaries and visualizations.
+@sec-model describes the three machine learning models being used in our study.
+@sec-results explores key findings from our analysis.
+Lastly, @sec-discussion addresses the limitations of the analysis and offers recommendations for reducing delays and improving subway service reliability.
+
 # Data {#sec-data}
 
-The dataset used here is downloaded from the IPUMS USA website via @ipums_usa_v15.
+The dataset used here is downloaded from the IPUMS USA website via [@ipums_usa_v15].
 Our project's extracted dataset contains all the responses from the American Community Survey (ACS) from 2023.
 Our extraction included the following variables: total income, dwelling mortgage status, sex of the individual, marital status, education attainment, private or public school, industry or occupation, veteran status, and age.
-We acknowledge that some of the variables (such as educational level, occupation, etc.) we have chosen here may have a more detailed version that contains more information.
+Some of the variables (such as educational level, occupation) we have chosen here have a more detailed version that contains more information.
 However, due to the nature of categorical variables, having detailed variables will create more classes of that variable, thus leading to more features for our model when encoding those variables.
 Although more features will improve the overall performance of our predictive model, it will also create issues such as overfitting, needing more computational resources, and lack of interpretability.
-Due to the trade-off of these mentioned pros and cons, we have decided to proceed with these higher-level variables for this project for simplicity.
+We have chosen higher-level variables in this study for better computational efficiency.
 Furthermore, we have cleaned the data for better data quality.
@@
 
-## Mortgage
+## Descriptive Data Analysis
+
+In this subsection, we will explore every variable used in our study by examining some of the basic statistics. These preliminary analyses can help us better understand the data we are working with.
+
+### Mortgage
 The Mortgage variable describes the mortgage status of dwelling.
@@
 #| message: false
-#| fig-pos: "H"
-#| fig-cap: Mortgage Status Distribution
+#| fig-pos: "htp"
+#| fig-cap: Most Homeowners Are on Mortgage
 
@@
 data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
+mortgage_labels = {
+    '1': 'No, owned free and clear',
+    '3': 'Yes, mortgaged',
+    '4': 'Yes, contract to purchase'
+}
+data['MORTGAGE'] = data['MORTGAGE'].astype(str).map(mortgage_labels)
 plt.figure(figsize=(5, 3))
@@
 
-## Sex
+### Sex
 This variable is coded as a binary variable, and we will use it as-is in our project.
@@
 The binary coding is a practical choice to maintain consistency with the dataset, though it may not fully capture the complexity and diversity of gender identities.
-@fig-figure2 below shows the distribution of this variable.
+@tbl-table1 shows the distribution of this variable.
 
@@
 #| echo: false
-#| label: fig-figure2
+#| label: tbl-table1
 #| warning: false
 #| message: false
-#| fig-pos: "H"
-#| fig-cap: Sex Distribution
+#| fig-pos: "htp"
 
@@
 import matplotlib.pyplot as plt
-import seaborn as sns
+
 data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
-plt.figure(figsize=(5, 3))
-sns.countplot(x='SEX', data=data, palette='pastel', edgecolor='black')
-plt.xlabel('Sex', fontsize=12)
-plt.ylabel('Count', fontsize=12)
-plt.xticks(ha='right')
-plt.grid(axis='y', linestyle='--', alpha=0.7)
+sex_labels = {'1': 'Male', '2': 'Female'}
+data['SEX'] = data['SEX'].map(sex_labels)
+vet_status_count = data['SEX'].value_counts().reset_index()
+vet_status_count.columns = ['Sex', 'Count']
+fig, ax = plt.subplots(figsize=(6, 2))
+ax.axis('off')
+table = ax.table(
+    cellText=vet_status_count.values,
+    colLabels=vet_status_count.columns,
+    cellLoc='center',
+    loc='center'
+)
+table.auto_set_font_size(False)
+table.set_fontsize(10)
+table.scale(1, 1.5)
 plt.tight_layout()
@@
 
-## Age
+### Age
 Age describes the age of the individual, and a age of 0 was coded for the individuals under 1 year old.
 For our dataset, the minimum age is 16, however.
-Age distribution of our dataset is shown below in @fig-figure3.
+Age distribution of our dataset is shown below in @fig-figure2.
 ```{python}
@@
 #| echo: false
-#| label: fig-figure3
+#| label: fig-figure2
 #| warning: false
@@
 
-## Marital Status
+### Marital Status
 Marital status describes the marital status of the individual.
@@
 Details available in @sec-datacleaning.
-The following @fig-figure4 illustrate the distribution of marital status of the participants in our dataset.
+The following @fig-figure3 illustrate the distribution of marital status of the participants in our dataset.
 ```{python}
@@
 #| echo: false
-#| label: fig-figure4
+#| label: fig-figure3
 #| warning: false
@@
 #| fig-pos: "H"
-#| fig-cap: Counts of Marital Status
+#| fig-cap: Counts of Marital Status, Most Participants Are Married with Spouse Present, or Never Married/Single
 
@@
 }
+order = [
+    'Married, spouse present',
+    'Married, spouse absent',
+    'Separated',
+    'Divorced',
+    'Widowed',
+    'Never married/single'
+]
 data['MARST'] = data['MARST'].map(marrital_status_label)
 plt.figure(figsize=(10, 6))
-sns.countplot(x='MARST', data=data, palette='pastel', edgecolor='black')
+sns.countplot(x='MARST', data=data, order=order, palette='pastel', edgecolor='black')
 plt.xlabel('Marital Status', fontsize=12)
@@
 
-## Educational Attainment [general version]
+### Educational Attainment
 This is a general version (compared to the detailed version).
@@
 This column represents the highest educational attainment achieved by the individual.
-The distribution of the variable is shown in @fig-figure5.
+The distribution of the variable is shown in @fig-figure4.
 ```{python}
@@
 #| echo: false
-#| label: fig-figure5
+#| label: fig-figure4
 #| warning: false
@@
 #| fig-pos: "H"
-#| fig-cap: Educational Attainment
+#| fig-cap: Most Participants Have Middel School Attainment or Above
 
@@
 }
+order = [
+    'No Schooling',
+    'Nursery School',
+    'Primary School',
+    'Middle School',
+    'College 1-4 Years',
+    'College 5+ Years'
+]
 data['EDUC_new'] = data['EDUC_new'].map(educ_new_label)
 plt.figure(figsize=(10, 6))
-sns.countplot(x='EDUC_new', data=data, palette='pastel', edgecolor='black')
+sns.countplot(x='EDUC_new', data=data, order=order, palette='pastel', edgecolor='black')
 plt.xlabel('Educational Attainment', fontsize=12)
@@
 
-## School Type
+### School Type
 School type indicates if the individual went to a public school or private school.
@@
 We are not sure how the participants would respond if they have been to more than one type of schools.
-@fig-figure6 shows the distribution of our sample.
+@fig-figure5 shows the distribution of our sample.
 ```{python}
@@
 #| echo: false
-#| label: fig-figure6
+#| label: fig-figure5
 #| warning: false
@@
 ```
-## Industry
+
+### Industry
 Industry describes the type of industry the individual works in.
@@
 
-## Veteran Status [General Version]
+### Veteran Status
 Veteran status idicates whether the individual is a veteran or not.
@@
 There is also a detailed version of this variable, which is not used in our project.
-The counts is displayed in @tbl-table1.
+The counts is displayed in @tbl-table2.
 ```{python}
@@
 #| echo: false
-#| label: tbl-table1
+#| label: tbl-table2
 #| warning: false
@@
 #| fig-pos: "H"
-#| fig-cap: Veteran Status
 
@@
 
-## Total Income
+### Total Income
 Total income is the total income made by the individual for the year.
-The distribution of total income is depicted in @fig-figure7.
+The distribution of total income is depicted in @fig-figure6.
 We have removed the outliers from the graph to have a better understanding of the general pattern of the overall distribution.
+From the figure, it is clear that more income are reported near multiple of $10,000 as a unit.
+This may be caused by the "heaping and leaping" phenomonon found in self-reported survey data [@pudney2008heaping].
 ```{python}
@@
 #| echo: false
-#| label: fig-figure7
+#| label: fig-figure6
 #| warning: false
@@
 ```
-# Exploratory Data Analysis {#sec-eda}
 
-## Income Distribution by Mortgage Status
+## Analysis of Feature Relationships {#sec-afr}
+
+In this subsection, we will explore some of the potential relationships among the features chosen for our study.
+Understanding these relationships can prepare us better for our model construction in the next section.
+
+### Income Distribution by Mortgage Status
 
@@
 
-## Correlation Analysis of Age, Education, and Income
+### Correlation Analysis of Age, Education, and Income
 
@@
 
-## Average Income by Education and Gender {#sec-eda3}
+### Average Income by Education and Gender {#sec-afr3}
 \begin{figure}[H]
@@
 We implemented three models: Linear Regression with Interaction Terms, Random Forest, and XGBoost, each chosen for their unique strengths in prediction and interpretation.
+
+## Train Test Split of Data
+
+We have cleaned our data before we split it into training and testing datasets.
+We are aware of potential data leakage of our testing data, and we have mitigated this risk by not using overall statistics or normalization that requires the original data to calculate, not imputing any missing data, or engineering features that include the original dataset.
+By doing so, we can minimize the risk of data leakage, and ensure the perfect seventy and thirty percent train test split.
 
@@
 
-Note that according to our previous analysis in @sec-eda3, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
+Note that according to our previous analysis in @sec-afr3, the interaction term, `EDUC_SEX_INTERACTION`, was included to capture the combined effect of education level and gender on income.
 This allowed us to model the non-linear relationship between these two variables and their joint impact on the outcome variable.
@@
 |--------------------------|-----------|
-| cat__STATEICP            | 3.242254  |
-| cat__GQ                  | 1.002419  |
-| cat__MORTGAGE            | 2.658013  |
-| cat__MARST               | 1.741158  |
-| cat__SCHLTYPE            | 1.239412  |
-| cat__OCC2010             | 3.187731  |
-| cat__IND1990             | 8.490997  |
-| num__AGE                 | 7.871046  |
-| num__EDUC_SEX_INTERACTION| 6.945850  |
+| cat__STATEICP            | 3.24  |
+| cat__GQ                  | 1.00  |
+| cat__MORTGAGE            | 2.66  |
+| cat__MARST               | 1.74  |
+| cat__SCHLTYPE            | 1.24  |
+| cat__OCC2010             | 3.19  |
+| cat__IND1990             | 8.49  |
+| num__AGE                 | 7.87  |
+| num__EDUC_SEX_INTERACTION| 6.95  |
 :::
@@
 |----------------------------|--------------------|
-| R-squared                  | 0.0774938018       |
-| Adjusted R-squared         | 0.0774637241       |
+| R-squared                  | 0.0774       |
+| Adjusted R-squared         | 0.0774     |
 | Mean Squared Error (MSE)   | 7,448,800,507.01   |
@@
 |-----------------|------------|
-| cat\_\_EDUC_new | 0.337065   |
-| cat\_\_IND1990  | 0.243454   |
-| num\_\_AGE      | 0.193926   |
-| cat\_\_SEX      | 0.149342   |
-| cat\_\_MARST    | 0.055948   |
-| cat\_\_MORTGAGE | 0.013064   |
-| cat\_\_SCHLTYPE | 0.004251   |
-| cat\_\_VETSTAT  | 0.002948   |
+| cat\_\_EDUC_new | 0.337   |
+| cat\_\_IND1990  | 0.243   |
+| num\_\_AGE      | 0.194   |
+| cat\_\_SEX      | 0.149   |
+| cat\_\_MARST    | 0.056   |
+| cat\_\_MORTGAGE | 0.013   |
+| cat\_\_SCHLTYPE | 0.004   |
+| cat\_\_VETSTAT  | 0.003   |
 :::
@@
 We performed cross-validation using cross_val_score function from scikit-learn  [@cite_sklearn] to evaluate the robustness of the model.
+
+\newpage
 
@@
 
-One consistent finding across all models is the significant impact of education and industry on income.
-These factors were consistently ranked as the most important predictors in both Random Forest and XGBoost models.
-Age also played an important role but to a lesser extent, while variables such as marital status and sex showed moderate influence.
+### Feature Importance
+We have found that educational attainment and occupation industry are consistently the most important features in all three models, especially in XGBoost and Random Forest.
+This indicates that higher educational attainment and certain industries are correlated with higher personal income.
+This confirms our original hypothesis that certain demographic factors, such as education, may lead to specialized industry sectors where more income can be generated.
 
-The residual analysis revealed that none of the models perfectly fit the data, particularly for higher income levels.
-Both Random Forest and XGBoost exhibited a pattern of increasing residual variance (heteroscedasticity) as predicted income increased, which suggests that the models struggled to accurately predict high-income values.
-Linear Regression, while more interpretable, performed the worst in terms of accuracy, especially for non-linear relationships between the predictors and income.
+Age is also a useful predictor.
+However, despite the common belief that age is a determining factor when predicting personal income, its effectiveness is not as strong as education or industry.
+This may be caused by the fact that in some industries, your experience may hit a plateau, after which age may not contribute as much to experience as before certain age.
 
-Despite the overall success of XGBoost, it is clear that there are areas where the model's performance could be further refined.
+Marital status and sex do affect personal income, but not as significantly as the other features.
+It is great to prove that certain demographical features are not as informative when predicting personal income, suggesting a "fair" distribution of wealth.
+
+### Models
+
+None of the three models used in our study provided an excellent prediction of personal income.
+For the residual analysis, it appears that the models' predictbility deterioate as the self-reported income gets higher.
+REsidual variance increased as the predicted income increased in the Random Forest and XGBoost model, suggesting potential heteroscedasticity.
+Our educated guess is that extra-high income may come from various sources, and simplifying them into a simple model like this is not a reasonable solution.
+
+Among the three models, XGBoost has the highest accuracy overall.
+We believe that it can capture more complex relationships and interactions among the variables than the other two models.
+However, the high MSE for extra-high income individuals indicates that the model still needs to be refined more to achieve an accurate prediction.
+
+Linear Regression has underperformed relative to the other two models.
+We speculate that simplicity in modeling linear relationships did not provide sufficient information to capture the non-linear patterns underlying our data, especially where multiple features interact to influence income, as suggested in @sec-afr.
 

### 2024-11-18 paper/paper.qmd
Revise Discussion and Render Paper

@@
 However, this does not mean it is without limitations or that further improvements cannot be made.
+This paper aims to analyze the relationship between demographic and categorical factors and personal income through predictive modeling using Linear Regression, Random Forest, and XGBoost.
 
@@
 ### Feature Importance
+
 We have found that educational attainment and occupation industry are consistently the most important features in all three models, especially in XGBoost and Random Forest.
@@
 This may be caused by the fact that in some industries, your experience may hit a plateau, after which age may not contribute as much to experience as before certain age.
+It would be interesting for future studies to explore non-linear relationships between age and income by incorporating interaction terms, such as age and education or age and industry.
 
@@
 It is great to prove that certain demographical features are not as informative when predicting personal income, suggesting a "fair" distribution of wealth.
+However, the subtle influence of gender raises questions about income equity.
+While gender was less impactful overall, its specific effects in certain industries could be explored further.
 
@@
 For the residual analysis, it appears that the models' predictbility deterioate as the self-reported income gets higher.
-REsidual variance increased as the predicted income increased in the Random Forest and XGBoost model, suggesting potential heteroscedasticity.
+Residual variance increased as the predicted income increased in the Random Forest and XGBoost model, suggesting potential heteroscedasticity.
 Our educated guess is that extra-high income may come from various sources, and simplifying them into a simple model like this is not a reasonable solution.
@@
 However, the high MSE for extra-high income individuals indicates that the model still needs to be refined more to achieve an accurate prediction.
+Adding more predictors or considering advanced ensemble methods might help better account for the variability at higher income levels.
 
@@
 We speculate that simplicity in modeling linear relationships did not provide sufficient information to capture the non-linear patterns underlying our data, especially where multiple features interact to influence income, as suggested in @sec-afr.
+It serves as a useful baseline but is less practical for this complex data structure.
 
@@
 
+Additionally, the use of cross-sectional data in this analysis limits our ability to draw causal inferences.
+While the models identify associations, they do not account for temporal or dynamic effects, such as career progression, economic cycles, or policy changes, which might influence income over time.
+Longitudinal data could provide more robust insights into these dynamics.
+
 ## Next Steps
@@
 For example, interactions between age and education, or industry and gender, could capture more nuanced relationships.
+Adding geographic location or labor market data could also provide deeper insights into regional income disparities.
 
@@
 
--   The skewness in income distribution poses a significant challenge for modeling.
-Exploring techniques like synthetic data generation for high-income individuals or applying weighting schemes to account for data imbalance might help the models better capture the variance across all income levels.
+-   Future studies could explore the temporal dimension of income predictors by incorporating longitudinal data, enabling an analysis of income trajectories over time.
+This would help assess not only predictors but also trends and changes in income inequality.
 
@@
 Future work will focus on addressing these issues while exploring more advanced modeling techniques and feature engineering to enhance predictive performance further.
+A deeper dive into causal relationships and longitudinal data could also uncover additional factors influencing income dynamics.
 
